Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-05-14 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Added edits to Conclusions too.

\pagebreak
\section{Conclusions}
The atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known despite of vast amount of accurate experimental
data. Atomistic resolution molecular dynamics simulation model which reproduced the
experimental data would automatically resolve the structures giving an interpretation of experimental results.
In this work we have collected and reviewed the experimental C--H bond vector order
parameters available in the literature. These experimental parameters are then compared to
different atomistic resolution simulation models for a fully hydrated bilayer, bilayers dehydrated to different extents, and
lipid bilayers containing various amounts of cholesterol. Our results have led to the following conclusions:

- The C--H bond order parameters measured with different NMR techniques are in good agreement
with each others. By combining the experimental results from various sources we concluded
that the order parameters for each C--H bond are known with a quantitative accuracy of $\pm$0.02.

- Comparison of order parameters between experiments and different atomistic resolution models 
together with the structural analysis shows that these parameters can be used
to judge the structural accuracy of the model. Thus the combination of atomistic resolution 
molecular dynamics simulations and NMR experiments can be used to resolve the atomistic resolution
structures of biomolecules in biologically relevant conditions. 
This approach can be extended from lipids also, for example, to membrane proteins.

@@

- The review of previous experimental results revealed that the 
choline order parameters are increasing when the bilayer is dehydrated. Our simulations suggest that
this can be explained by P--N vector tilting more parallel to the membrane. 
These results strongly supports and complements the idea that the charge induced choline tilting and 
partition can be measured by using the order parameter changes~\cite{ionpaper,scherer89}.

@@

- Only modest changes of glycerol backbone and choline order parameters are observed experimentally with increasing cholesterol concentration.
When interpreted by using the simulations with the most realistic response to the cholesterol concentration,
this observation can be explained such that the cholesterol induces only minor changes in the g$_3$-g$_2$-g$_1$-O(sn-1)
dihedral in the glycerol backbone.

@@

- Besides the main conclusions, we have created the most extensive
publicly available collection of molecular dynamics simulation trajectories of lipid bilayers
into the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids}. 
This collection opens up numerous possiblities for different analyses with
much less effort than previously required.

In general, we conclude that to fully utilize the potential of atomistic resolution classical molecular dynamics simulations
in the structural interpretation of the high resolution NMR data~\cite{ferreira14} for lipid bilayers one has to  
improve the phoshatidylcholine glycerol backbone and choline headgroup parameters.

This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an communication
platform. All the scientific contributions have been done through the blog and are publicly
available~\cite{nmrlipids}.

>>>>

\pagebreak
\section{Conclusions}
The atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known despite of vast amount of accurate experimental
data. An atomistic resolution molecular dynamics simulation model that would reproduce the
experimental data would automatically resolve the structures, thus giving an unprecedently detailed interpretation of the experimental data.
In this work we have collected and reviewed the experimental C--H bond vector order
parameters available in the literature. These accurate experimental data are then compared to 13
different atomistic resolution simulation models for a fully hydrated lipid bilayer system, followed by bilayers dehydrated to different extents, and
finally bilayers containing various amounts of cholesterol. We are led to the following four main conclusions:

(1) The C--H bond order parameters measured with different NMR techniques are consistent.
By combining the experimental results from various sources we concluded
that the order parameters for each C--H bond are known with a quantitative accuracy of $\pm$0.02.

(2) Comparison of order parameters between experiments and different atomistic resolution models 
together with structural analysis showed that the order parameters can be used
to judge the structural accuracy of a model. Thus the combination of atomistic resolution 
molecular dynamics simulations and NMR experiments can be used to resolve the atomistic resolution
structures of biomolecules in biologically relevant conditions. 
This approach can be extended from lipids to, for example, membrane proteins.

@@

(3) The review of previous experimental results revealed that when a bilayer is dehydrated the 
choline order parameters increase. Our simulations suggested that
this can be explained by the P--N vector tilting more parallel to the membrane. 
This strongly supports and complements the idea that charge-induced choline tilting
can be measured using order parameter changes~\cite{ionpaper,scherer89}.

@@

(4) Only modest changes of glycerol backbone and choline order parameters are observed experimentally with increasing cholesterol content.
When interpreted using the computational lipid model that we found to have the most realistic response to cholesterol,
this observation means that cholesterol induces only minor changes in (the g$_3$-g$_2$-g$_1$-O(sn-1)
dihedral of) the glycerol backbone, in other words, there is no major conformational change of the lipid.

@@

(+) Besides these four main conclusions, we note that we have created the most extensive
publicly available collection of molecular dynamics simulation trajectories of lipid bilayers
(\url{https://zenodo.org/collection/user-nmrlipids}). The mere existence of
this collection opens up numerous possibilities for unforeseen analyses, such as data mining,
and rapid testing of ideas with
much less computational effort than previously required.

In general, we conclude that in order to fully utilize the potential of atomistic-resolution classical molecular dynamics simulations
in the structural interpretation of high resolution NMR data~\cite{ferreira14} for lipid bilayers, one must  
improve the phoshatidylcholine glycerol backbone and choline headgroup parameters of the existing force fields.

This work has been done as a fully open collaboration, using \url{nmrlipids.blogspot.fi} as the communication
platform. All the scientific contributions have been communicated publicly through this blog~\cite{nmrlipids}.
